---
title: "Model"
subtitle: "What is the difference between black, white, and hispanic wages through each education level?"
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| message: false
#| warning: false
library(tidyverse)
library(primer.data)
library(brms)
library(tidybayes)
library(gtsummary)
library(dplyr)
```

```{r}
#| cache: true
large_data <- read_csv("C:/Users/Satis/projects/average-wages/wages_by_education.csv")
```

To answer the question of the difference in wages among black, white, and hispanic workers across different education levels, we first analyze the wage data using a Bayesian regression model. For this initial model, we use the following formula:

$$
\text{wage} = \beta_{0} + \beta_{1} \times \text{education}_i
+ \beta_{2} \times \text{year}_i
$$

```{r}
#| cache: true

fit_education <- brm(formula = year ~ less_than_hs + high_school + some_college + bachelors_degree + advanced_degree,
                  data = large_data,
                  family = gaussian(),
                  refresh = 0,
                  silent = 2,
                  seed = 9)

tbl_regression(fit_education)
```

We observe the impact of different education levels on average hourly wages. Next, we calculate the effect of education level and ethnicity on wages using the following model:

$$
\text{wage} = \beta_{0} + \beta_{1} \times \text{education}_i + \beta_{2} \times \text{ethnicity}_i 
$$
```{r}
#| cache: true

fit_education_year <- brm(formula = high_school ~ white_less_than_hs + white_high_school + white_some_college +
                            white_bachelors_degree + white_advanced_degree + black_less_than_hs +
                            black_high_school + black_some_college + black_bachelors_degree + black_advanced_degree +
                            hispanic_less_than_hs + hispanic_high_school + hispanic_some_college + hispanic_bachelors_degree +
                            hispanic_advanced_degree,
                  data = large_data,
                  family = gaussian(),
                  refresh = 0,
                  silent = 2,
                  seed = 9)

fit_education_year
```

This model includes the ethnicity variable to account for differences between black, white, and hispanic workers, providing a more comprehensive analysis.

```{r}
tbl_regression(fit_education_year)
```

In summary, we find that higher education levels generally correspond to higher wages. However, there are significant differences in wages among black, white, and Hispanic workers at each education level. This relationship highlights the importance of considering both education and ethnicity when analyzing wage disparities in the USA.
